---
title               : "Flex Knit"
author              : "Han Oostdijk"
date                : "file created : `r format(Sys.time(), '%d%b%Y')`"
output              :
  pdf_document      :
    keep_tex        : yes
  # html_document     :
  #   keep_md         : yes
classoption         : portrait # orlandscape 
# do not place blanks between geometry and ':' or the option will be ignored !
geometry: [
  left=0.6in ,
  right=0.5in ,
  top=1in ,
  bottom=1in
  ]
header-includes     : [
  '\usepackage[scaled]{DejaVuSansMono}' ,
  '\renewcommand*{\familydefault}{\sfdefault}' 
  ]
urlcolor            : blue
hoqc_version        : _v1
# hoqc_ filenames can have a path
hoqc_output         : './output/Flex Knit' 
hoqc_yaml           : ./output/yaml
hoqc_yaml_new       : ./output/yaml_new
hoqc_copy           : ./output/flexknit.Rmd
hoqc_force_ext      : yes
knit                : (function (...) { source('myknit.R'); myknit(...) })
# do not place blanks between params and ':' !
params:
  parm1             : myfirstparm
---

## Synopsis
This document presents the R-function `myknit` that enables dynamically changing the output files of the knit-process by using the undocumented `knit` yaml option.

## Acknowledgement
The idea for manipulating yaml information in the way done here was found in a [Stack Overflow  article](https://stackoverflow.com/questions/39885363/importing-common-yaml-in-rstudio-knitr-document) (option 2) authored by *mathematical-coffee*.

## Introduction
When creating a new document I always have to try various versions of the yaml metadata block before I have a version that works for me. I need an easy way to keep these various inputs and outputs separated. The standard way (the `knit` button) always creates/overwrites the same output file. The `rmarkdown::render` function that is called as a result of clicking the button allows specifying a name for the output file. I found (see acknowledgement) a method to call this function with the undocumented (why?) `knit` yaml option. The `knit` option option specifies what is executed when the `knit` button is clicked: a direct call to `rmarkdown::render` or a user function.  
I use the construction `knit:  (function (...) { source('myknit.r'); myknit(...) })` with the user written function `myknit`. The `myknit` function supports some additional options in the yaml metadata block and uses these to construct output names.  

## Functionality of the `myknit` function
By specifying additional options in the yaml metadata block the user can do one or more of the following things:

- specify an alternative name for the output file
- append a version indicator to the name of the output file
- force that the proper extension is given to the name of the output file (when not specified)
- create a file with the yaml metadata block that was specified with a name including the version indicator
- create a file with the yaml metadata block after processing with a name including the version indicator
- create a copy of the input file with a name including the version indicator

## How to use the functionality of the `myknit` function
The `myknit` function is controlled by additional options in the yaml metadata block. To distinguish these options from the regular ones, they are prefixed with `hoqc_`. The additional options (with defaults in parentheses) are:  

- `hoqc_output    ('')` : the name of the output document file
- `hoc_version    ('')` : suffix to be given to files to indicate version
- `hoqc_force_ext  (F)` : should a filename extension be forced for document and yaml files? (`TRUE` or `FALSE`)
- `hoqc_yaml      ('')` : the name of file containing the original yaml metadata block
- `hoqc_yaml_new  ('')` : the name of file containing the new (modified) yaml metadata block
- `hoqc_copy      ('')` : the name of file containing a copy of the input file

The output document will in principle be named `hoqc_output` if this is specified; otherwise it is named after the input file. If `hoqc_output` is specified without an extension and `hoqc_force_ext == T` then the output name will be given an extension of `pdf` or `html` depending on the (first encountered, not commented) document_type. `hoc_version` will be inserted after the basename of the file. 

The orginal yaml metadata block will be written to an output file with name `hoqc_yaml` if this is specified.
After reading the orginal yaml metadata block comment lines (beginning with `#`) are removed. 
The modified yaml metadata block will be written to an output file with name `hoqc_yaml_new` if this is specified. In both yaml file names the `hoc_version` will be inserted and a `txt` extension will be given when an extension was not specified and `hoqc_force_ext == T`.  

A copy of the input file is created with name `hoqc_copy` if this is specified. The `hoc_version` will be inserted and a `Rmd` extension will be given when an extension was not specified and `hoqc_force_ext == T`. 

In the modified metadata block the additional options are included (added if it already exists) in the `params` block. The name of the output files are expanded to the full path. The next paragraph shows how the `params` block can be used to include e.g. the yaml metadata block in the document by specifying `r` `params$hoqc_yaml` within backticks. 

A usage scenario for working with LaTeX could be:  

- set `hoqc_output` and `hoqc_copy` e.g. to `./output/flexkit` and `keep_tex` to `yes` 
- set `hoc_version` to `_v1` and knit the file. 
- set `hoc_version` to `_v2`, make some changes and knit again 

Intermediate folders will be created if necessary.

In the `output` subfolder one now can find for both versions the saved inputs (`flexknit_v1.Rmd` and `flexknit_v2.Rmd`) and the corresponding `tex` files and `pdf` files (with version indicators). When something is not working as expected one can compare the various versions to see what is wrong. In this way I saw that the yaml options for `geometry` and `params` do not allow blanks between the options and the colon.

## Listing of the yaml metadata blocks
As an example the original yaml (in `r params$hoqc_yaml`) for this document will be shown:
```{r echo=T,results='markup',comment='   ' }
cat(paste0(readLines(params$hoqc_yaml),collapse = '\n'))
```

and also the modified yaml (in `r params$hoqc_yaml_new`) :
```{r echo=T,results='markup',comment='   ' }
cat(paste0(readLines(params$hoqc_yaml_new),collapse = '\n'))
```

[//1]: # (This is a comment that is not shown in the output, as is the next section)
[//2]: # (see https://stackoverflow.com/questions/4823468/comments-in-markdown)

<!---

## Listing of the user-written function *myknit* and auxiliary functios

```{r echo=F,eval=F,results='markup',comment='' }
cat(paste0(readLines('myknit.R'),collapse = '\n'))
```
-->


## Session Info
```{r echo=F,results='latex',comment='   ' }
sessionInfo()
```

