---
params:
  doc_version: x.y
  altplot: !r FALSE
title: "`r paste0('iris_data_set_vm4 v',params$doc_version)`"  
author: "Han Oostdijk (www.hanoostdijk.nl)"
date: "`r format(Sys.time(), '%d %B, %Y')`" # current date
graphics: yes
linkcolor: blue
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: 
      - styles.tex
      - extra1.tex
      before_body: extra2.tex
    keep_tex: no 
    highlight: tango 
geometry: 
  - a4paper 
  - landscape 
toc: true
toc_depth: 2
knit:  (function(inputFile, encoding) {rmarkdown::render(inputFile,encoding=encoding,clean=T,output_file = 'Iris_data_set_vm4.pdf') })
bibliography: mybib.bib 
---

```{r child='iris_data_set_bib1.rmd'}
```

# Load data and create plot object
Before plotting the iris data set (in `r ref_tab('r1a',F)`) we list the first `r numlist` (because we set variable \mytextit{numlist} to `r numlist` in a chunck we do not present to the reader) observations in the data set in  `r ref_tab('lbltab1',T)`.  
As an example of a reference we use [@myrefa]. You can also include a reference in an inline footnote as demonstrated here^[This footnote references [@myrefb] and [@article].].   
NB. using the \LaTeX \ \mytextbf{\BSLASH{footnote}}-command does not give the correct layout.

```{r child='iris_data_set_bib2.rmd'}
```

```{r echo=F,results='asis'}
# to include References in the table of contents
#cat('\\cleardoublepage')
#cat('\\phantomsection')
cat('\\addcontentsline{toc}{section}{References}')
```

# References
