---
params:
  doc_version: x.y
  altplot: !r TRUE
title: "`r paste0('iris_data_set_vm3 v',params$doc_version)`"  
author: "Han Oostdijk (www.hanoostdijk.nl)"
date: "`r format(Sys.time(), '%d %B, %Y')`" # current date
graphics: yes
linkcolor: blue
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: 
      - styles.tex
      - extra1.tex
      before_body: extra2.tex
    keep_tex: no 
    highlight: tango 
geometry: 
  - a4paper 
  - landscape 
toc: true
toc_depth: 2
knit:  (function(inputFile, encoding) {rmarkdown::render(inputFile,encoding=encoding,clean=T,output_file = 'Iris_data_set_vm3.pdf') })
---

```{r child='iris_data_set_bib1.rmd'}
```

# Load data and create plot object
Before plotting the iris data set ( `r ref_tab('r1a',F)`) we list the first `r numlist` 
(because we set variable \mytextit{numlist} to `r numlist` in a chunck we do not present to the reader) 
observations in the data set `r ref_tab('lbltab1',T)`.  
As an example of a reference we use \cite{myrefa} with another reference in a footnote\footnote{see \cite{myrefb}}.
NB. we now have to use the \LaTeX \  \mytextbf{\BSLASH{cite}} and \mytextbf{\BSLASH{footnote}} commands explicitly.

```{r child='iris_data_set_bib2.rmd'}
```

```{r echo=F,results='asis'}
# to include References in the table of contents
#cat('\\cleardoublepage')
#cat('\\phantomsection')
cat('\\addcontentsline{toc}{section}{References}')
```

\begin{thebibliography}{20}

\bibitem[author1]{myrefa} example of reference for author1

\bibitem[author2]{myrefb} example of reference for author2 

\end{thebibliography}
