---
title: "iris_data_set_vm2 (minimal dependency on LaTeX)"  
author: "Han Oostdijk (www.hanoostdijk.nl)"
date: "December 2015"
graphics: yes
output: pdf_document
---

```{r child='setup.rmd'}
```

# Purpose of this example
In 'iris_data_set_vm1.rmd', where we did not use Latex, we could not use references to figures and tables.
Here we show how this becomes available when we use Latex (but obviously at the cost of introducing differences in the output formats).  
The only difference is that we assign labels and captions to the table (the function *def_tab* that is used in *kable* sets the label and caption) and the figure (with the chunk name as label and the contents of *fig.cap* as caption). With the function *ref_tab* we use these labels. *def_tab* and *ref_tab* are defined in the setup chunk.

# Libraries used

```{r echo=TRUE,warning=FALSE,message=FALSE}
library(ggplot2)
library(ggthemes)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
# constants/calculations we do not show in the text
numlist  =   10 # number of observations in iris data set that is listed
tabcap1  =   sprintf('first %.0f observations of iris data set',numlist)
figr1a   =   "default theme"
```

# Load data
We load the iris data set in the workspace.

```{r}
data(iris)
```

# List the part of the iris data set
We list the first `r numlist` 
(because we set variable *numlist* to `r numlist` in a chunck we do not present to the reader) 
observations in the data set `r ref_tab('lbltab1',T)`. 

```{r results='asis'}
kable(iris[1:numlist,],row.names=F,longtable=F,caption=def_tab('lbltab1',tabcap1))
cat(paste(' #produced',ref_tab('lbltab1',T,prefix='')))
```

# Create a plot of the iris data set
Make a *ggplot2* plot object *p* of two variables in the iris data set 
and use it to create a plot `r ref_tab('r1a',F)`.

```{r r1a,fig.cap=figr1a,out.width="5in",out.height="5in",results='asis'}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) + 
  geom_point() 
p + labs(title = 'default theme')
cat(paste(' #produced',ref_tab('r1a','F',prefix='')))
```

# Session Info
```{r}
sessionInfo()
```
