---
title: "iris_data_set_vm7 (ready for html and PDF)"  
author: "Han Oostdijk (www.hanoostdijk.nl)"
date: "`r format(Sys.time(), '%d%b%Y')`"
output:
  html_document:
    keep_md: yes
  bookdown::pdf_document2:
    keep_tex: yes
    toc: no
    includes:
      in_header: 
      - extra1.tex
params:
  myseed: '2017'
  numlist: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
out_type <- knitr::opts_knit$get("rmarkdown.pandoc.to")
```

# Functions that enable captions to figures and tables
 
This document contains two sets of functions in the child chunk *captions_html_latex.rmd*. One set is used for html documents and the other for pdf documents. For pdf documents it does not work with the output *pdf_document* so it is necessary to use *bookdown::pdf_document2* . With the (default) Pandoc template for this output type a table of contents is produced. The table of contents can be omitted by specifying **toc:no** in the yaml header.

We use a child chunk because the code in it is also available for other documents.

```{r child="captions_html_latex.rmd"}
```

# Defines captions for plot and table

```{r }
init_caption('Figure','Table')           # indicate prefix for figures and tables
add_caption("fig1","f","my first plot")  # define a caption for the first (an only) figure
add_caption("tab1","t","my first table") # define a caption for the first (an only) table
```

# Libraries used

```{r echo=TRUE,warning=FALSE,message=FALSE}
library(knitr)
library(pander)
library(ggplot2)
library(ggthemes)
```

```{r echo=FALSE}
# constants/calculations we do not show in the text
numlist  = params$numlist # number of observations in iris data set that is listed
```

# Load data
We load the iris data set in the workspace.

```{r}
data(iris)
```

# List the part of the iris data set
We list the first `r numlist` (because we set variable *numlist* to `r numlist` in a chunck we do not present to the reader) observations in the data set. See  `r ref_caption("tab1","t")` `r ref_caption("tab1","t","page")` 

```{r results='asis',comment=""}
  pandoc.table(iris[1:numlist,],caption=tab_caption("tab1"))
```

# Create a plot of the iris data set
Make a *ggplot2* plot object *p* of two variables in the iris data set 
and use it to create a plot that is shown in `r ref_caption("fig1","f")` `r ref_caption("fig1","f","page")` 

```{r fig1,results='asis',fig.cap=fig_caption("fig1")}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) + 
  geom_point() 
p + labs(title = 'default theme')
```

# Session Info {#Session-Info-anchor}
```{r}
sessionInfo()
```

