---
params:
  sortorder: ascending
title: "iris_data_set_vm6"  
author: "Han Oostdijk (www.hanoostdijk.nl)"
date: "`r format(Sys.time(), '%d %B, %Y')`" # current date
graphics: yes
linkcolor: blue
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: 
      - styles.tex
    keep_tex: no 
    highlight: tango 
geometry: 
  - a4paper 
  - portrait 
  - margin=0.5in
toc: false
toc_depth: 2
knit:  (function(inputFile, encoding) {rmarkdown::render(inputFile,encoding=encoding,clean=T,output_file = 'Iris_data_set_vm6.pdf') })
#bibliography: mybib.bib 
---

```{r child='setup.rmd'}
```

# Introduction 
This example shows how to print two tables side by side. It makes use of the 
function \mytextit{print\_table\_sbs} that generates the \LaTeX \ commands that are necessary for this. If this is usable depends on the size of the tables and your eyesight.

# Utility functions and constants
```{r echo=T,warning=FALSE,message=FALSE}
numlist = 10 # number of observations in iris data set that is listed

print_table_sbs <- function (files,label,cap,caps,scalebox=0.85) {
# print xtables saved in files side-by-side
# derived from Marcin Kosiński
# http://stackoverflow.com/questions/23926671/side-by-side-xtables-in-rmarkdown
cat('\\begin{table}[ht]\n')
cat('\\centering\n')  
for (i in 1:length(files)) {
  tc  = caps[[i]]
  tci = paste0(label,letters[i])
  f   = files[[i]]
  c = '\\subfloat[%s]{\\label{table:%s}\\scalebox{%.3f}{\\input{./%s}}}\\quad\n'
  cat(sprintf(c,tc,tci,scalebox,f))
}
cat(sprintf('\\caption{%s}\n',cap))
cat(sprintf('\\label{table:%s}\n',label))
cat('\\end{table}') 
}
```

# Explanation
We will create two tables in R and use package \mytextbf{xtable} to write the \LaTeX \  representation to a tex-file. We use the iris data set as an example and will sort it in  ascending and descending order (by the variable \mytextit{Sepal.Length}). Because that is  twice (nearly) the same code we have created the function \mytextit{create\_xtable} to do this. In the function we select only the first/last `r numlist` observations.    
However the only important line is  \mytextit{print(xtable(. . . floating=FALSE)} .  
The function \mytextit{print\_table\_sbs} will take the two files and print them side by side  if that fits. Otherwise the second table will be positioned below the first one.
`r ref_tab('r1a',prefix='')` we will show the `r numlist` observations with the smallest and `r ref_tab('r1b',prefix='',add_page=F)` the ones with the greatest \mytextit{Sepal.Length}. 
```{r,warning=F,message=F}
library(xtable)
create_xtable<- function (decreasing,numlist,filename) {
  data(iris)
  tab = iris[order(iris$Sepal.Length,decreasing = decreasing),]
  print(xtable(tab[1:numlist,],row.names=F),
    file=paste0(filename,'.tex'),floating=FALSE)
}

create_xtable(F,numlist,filename='asc') 
create_xtable(T,numlist,filename='desc') 
```

```{r r1,results= 'asis'}
files = c('asc','desc') # filenames (without suffix tex)
label = 'r1' # label (sublabels have suffix 1,2, ...)
cap   = 'first/last observations of iris data set sorted by \\mytextit{Sepal.Length}'
caps  = c('first', 'last')
print_table_sbs(files,label,cap,caps,scalebox=0.5) 
cat(paste(' #produced',ref_tab('r1',prefix='')))
```


